---
title: "QA/QC Report"
date: "`r format(Sys.time(), '%Y-%m-%d %H:%M:%S')`"
output:
  pdf_document:
    toc: true
    toc_depth: 3
    number_sections: true
    latex_engine: pdflatex
header-includes:
    - \usepackage{titlesec}
    - \newcommand{\sectionbreak}{\clearpage}
params:
  input_csv: ""
  baseline_csv: ""
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  warning = FALSE,
  message = FALSE
)

library(dplyr)
library(knitr)
library(kableExtra)
library(ggplot2)
library(tidyr)
```

\newpage

```{r load-env, include=FALSE}
# Accept paths either from params (rmarkdown::render call) or environment vars
# (set by the SLURM wrapper via Sys.setenv before render)
input_csv    <- if (nchar(params$input_csv)    > 0) params$input_csv    else Sys.getenv("QAQC_INPUT_CSV")
baseline_csv <- if (nchar(params$baseline_csv) > 0) params$baseline_csv else Sys.getenv("QAQC_BASELINE_CSV")

stopifnot("QAQC_INPUT_CSV is not set"    = nchar(input_csv)    > 0)
stopifnot("QAQC_BASELINE_CSV is not set" = nchar(baseline_csv) > 0)

df       <- read.csv(input_csv,    stringsAsFactors = FALSE, check.names = FALSE)
baseline <- read.csv(baseline_csv, stringsAsFactors = FALSE, check.names = FALSE)
```
# Introduction

The following document provides a summary for the data linkage results presented in `r print(input_csv)`. The linkage results are compared against the input geocoding file provided by `r print(baseline_csv)` to evaluate the number of missing values, provide summary statistics, and ensure no data loss has occurred. 

For questions or concerns, please contact:
Jaime Hart: jaime.hart@channing.harvard.edu


This document was prepared `r Sys.info()[["user"]]` on `r Sys.Date()`

# Job Information
```{r job-info}
info <- data.frame(
  Field = c("Input CSV", "Baseline CSV", "R Version", "Rendered at",
            "SLURM Job ID", "Node"),
  Value = c(
    input_csv,
    baseline_csv,
    R.version.string,
    format(Sys.time(), "%Y-%m-%d %H:%M:%S %Z"),
    Sys.getenv("SLURM_JOB_ID",   unset = "N/A"),
    Sys.getenv("SLURMD_NODENAME", unset = "N/A")
  )
)
kable(info, booktabs = TRUE, col.names = c("Field", "Value"),
      caption = "Run metadata") %>% 
  kable_styling(latex_options = c("hold_position", "striped"), font_size = 9)
```

---

# Dataset Overview

```{r overview}
overview <- data.frame(
  Metric = c("Rows", "Columns", "Column Names"),
  Input = c(
    nrow(df),
    ncol(df),
    paste(colnames(df), collapse = ",\n")
  ),
  Baseline = c(
    nrow(baseline),
    ncol(baseline),
    paste(colnames(baseline), collapse = ", ")
  )
)
kbl(overview, 
    booktabs = TRUE, 
    col.names=c("Metric","Input","Baseline"),
    longtable=TRUE,
    caption = "Dataset dimensions",
    escape=TRUE) %>%   
  kable_styling(full_width=TRUE,
                latex_options = c("scale_down","hold_position", "striped","repeat_header"), 
                font_size = 9)
```

---

# Row Count Comparison

```{r row-comparison}
row_diff   <- nrow(df) - nrow(baseline)
row_status <- if (row_diff == 0) "PASS" else if (row_diff > 0) "WARNING – extra rows" else "WARNING – missing rows"

row_tbl <- data.frame(
  Metric      = c("Input rows", "Baseline rows", "Difference", "Status"),
  Value       = c(nrow(df), nrow(baseline), row_diff, row_status)
)
kbl(row_tbl, 
    booktabs = TRUE, 
    caption = "Row count QC",
    col.names=c("Metric","Value"),
    escape=TRUE) %>%
  kable_styling(latex_options = c("hold_position", "striped"), font_size = 9) %>%
  row_spec(4,
    bold      = TRUE,
    color     = ifelse(row_status == "PASS", "green", "red")
  )
```

---

# ID Column Comparison

```{r id-check}
has_id_input    <- "id" %in% colnames(df)
has_id_baseline <- "id" %in% colnames(baseline)

if (!has_id_input || !has_id_baseline) {
  cat("**WARNING:** 'ID' column not found in",
      if (!has_id_input)    "the input CSV"    else "",
      if (!has_id_baseline) "the baseline CSV" else "",
      "— ID comparison skipped.\n")
} else {
  ids_input    <- unique(df$id)
  ids_baseline <- unique(baseline$id)

  only_in_input    <- setdiff(ids_input,    ids_baseline)
  only_in_baseline <- setdiff(ids_baseline, ids_input)
  shared           <- intersect(ids_input,  ids_baseline)

  id_summary <- data.frame(
    Metric = c(
      "Unique IDs – input",
      "Unique IDs – baseline",
      "IDs in both",
      "IDs only in input (extra)",
      "IDs only in baseline (missing)",
      "Status"
    ),
    Value = c(
      length(ids_input),
      length(ids_baseline),
      length(shared),
      length(only_in_input),
      length(only_in_baseline),
      ifelse(length(only_in_input) == 0 & length(only_in_baseline) == 0,
             "PASS", "FAIL – ID mismatch")
    )
  )

  kbl(id_summary, 
      col.names=c("Metric","Value"),
      booktabs = TRUE, 
      caption = "ID column QC",
      escape=TRUE) %>%
    kable_styling(latex_options = c("hold_position", "striped"), font_size = 9) %>%
    row_spec(6,
      bold  = TRUE,
      color = ifelse(id_summary$Value[6] == "PASS", "green", "red")
    )
}
```

```{r id-detail, results='asis'}
if (exists("only_in_input") && length(only_in_input) > 0) {
  cat("\n## IDs present in input but NOT in baseline\n\n")
  show_n <- min(50, length(only_in_input))
  kbl(data.frame(ID = only_in_input[seq_len(show_n)]),
        booktabs = TRUE,
        caption  = sprintf("Extra IDs (showing %d of %d)", 
                           show_n, 
                           length(only_in_input)),
        escape=TRUE) %>%
    kable_styling(latex_options = c("hold_position", "striped"), font_size = 8) %>%
    print()
}

if (exists("only_in_baseline") && length(only_in_baseline) > 0) {
  cat("\n## IDs present in baseline but NOT in input\n\n")
  show_n <- min(50, length(only_in_baseline))
  kbl(data.frame(ID = only_in_baseline[seq_len(show_n)]),
        booktabs = TRUE,
        caption  = sprintf("Missing IDs (showing %d of %d)", 
                           show_n, 
                           length(only_in_baseline)),
        escape=TRUE) %>%
    kable_styling(latex_options = c("hold_position", "striped"), font_size = 8) %>%
    print()
}
```

---

# Summary Statistics — Input CSV

```{r summary-stats}
# Numeric columns
num_cols <- names(df)[sapply(df, is.numeric)]

if (length(num_cols) > 0) {
  num_summary <- df %>%
    select(all_of(num_cols)) %>%
    summarise(across(everything(), list(
      Min    = ~min(., na.rm = TRUE),
      Q1     = ~quantile(., 0.25, na.rm = TRUE),
      Median = ~median(., na.rm = TRUE),
      Mean   = ~mean(., na.rm = TRUE),
      Q3     = ~quantile(., 0.75, na.rm = TRUE),
      Max    = ~max(., na.rm = TRUE),
      SD     = ~sd(., na.rm = TRUE),
      NAs    = ~sum(is.na(.))
    ), .names = "{.col}__{.fn}")) %>%
    pivot_longer(everything(),
                 names_to  = c("Column", "Stat"),
                 names_sep = "__") %>%
    pivot_wider(names_from = "Stat", values_from = "value")

  kbl(num_summary, 
      booktabs = TRUE, 
      digits = 3,
      caption = "Numeric column summary statistics",
      col.names=c("Column","Min","Q1","Median","Mean","Q3","Max","SD","NAs"),
      longtable=TRUE,
      escape=TRUE) %>%
    kable_styling(latex_options = c("hold_position", "striped", "scale_down"),
                  font_size = 8)
} else {
  cat("No numeric columns detected.\n")
}
```

```{r char-summary, results='asis'}
# Character / factor columns — unique value counts
chr_cols <- names(df)[sapply(df, function(x) is.character(x) | is.factor(x))]
chr_cols <- chr_cols[chr_cols != "id"]  # ID handled separately above

if (length(chr_cols) > 0) {
  cat("\n## Categorical Column Summaries\n\n")
  for (col in chr_cols) {
    tbl <- df %>%
      group_by(Value = .data[[col]]) %>%
      summarise(Count = n(), .groups = "drop") %>%
      arrange(desc(Value)) %>%
      slice_head(n = 20)

    cat(sprintf("\n### %s (top %d of %d unique values)\n\n",
                col, nrow(tbl), n_distinct(df[[col]])))
    
    cleancaption<-knitr:::escape_latex(sprintf("'%s' value frequencies",col))
    
    kbl(tbl, 
        booktabs = TRUE,
        longtable=TRUE,
        caption = cleancaption,
        escape=TRUE,
        col.names=c("Value","Count")) %>%
      kable_styling(latex_options = c("hold_position", "striped"), font_size = 8) %>%
      print()
  }
}
```
---

# Missing Value Report

```{r missing}
missing_tbl <- data.frame(
  Column       = colnames(df),
  Missing      = sapply(df, function(x) sum(is.na(x))),
  Pct_Missing  = sapply(df, function(x) round(mean(is.na(x)) * 100, 2))
) %>%
  arrange(desc(Column))

kbl(missing_tbl, 
    booktabs = TRUE, 
    row.names = FALSE,
    col.names=c("Column","Missing","Percent Missing"),
      caption = "Missing values per column",
      escpape=TRUE) %>%
  kable_styling(latex_options = c("hold_position", "striped"), font_size = 9)
```

---

# QC Summary

```{r qc-final}
row_pass <- row_diff == 0
id_pass  <- if (exists("only_in_input")) {
  length(only_in_input) == 0 & length(only_in_baseline) == 0
} else { NA }

qc_tbl <- data.frame(
  Check  = c("Row count matches baseline", "All IDs match baseline"),
  Result = c(
    ifelse(row_pass, "PASS", "FAIL"),
    ifelse(is.na(id_pass), "SKIPPED", ifelse(id_pass, "PASS", "FAIL"))
  )
)

kbl(qc_tbl, 
    booktabs = TRUE, 
    caption = "Overall QC checklist",
    col.names=c("Check","Result"),
    escape=TRUE) %>%
  kable_styling(latex_options = c("hold_position", "striped")) %>%
  row_spec(which(qc_tbl$Result == "PASS"),    bold = TRUE, color = "green") %>%
  row_spec(which(qc_tbl$Result == "FAIL"),    bold = TRUE, color = "red") %>%
  row_spec(which(qc_tbl$Result == "SKIPPED"), bold = TRUE, color = "orange")
```
